---
title: "Model Selection for airdryer Data with Explanations"
output: 
  html_document: 
    highlight: textmate
    theme: cosmo
date: '2022-03-25'
---


First we need to upload the data. So:
```{r}
df_airdryer = read.csv('E:/2022/Freelance/Jobs/001-25/data-airdryer.csv', header=T)
df_airdryer = data.frame(time=df_airdryer$Time..hours., fails=df_airdryer$Number.of.failures)
```

So we created a new data frame with proper column names. To see what it looks like, we can print the data frame:
```{r}
df_airdryer
```

In the second column the number of failures for each time points is recorded. In order to have a complete data set we need to create repetitions of those time points with the frequency of the number of failures.
Therefore we repeat each time points relative to their respective number of failures.

```{r}
time_rep = rep(df_airdryer$time, df_airdryer$fails)
```

Now we create an empty data frame to hold the information of the distributions that we are going to check. 
```{r}
info_df = data.frame(model=NA, method=NA,Log_likelihood=NA, AIC=NA)
```

In this data frame we have a *model* column which stores the name of the distribution (weibull, gamma, log-normal, beta), a *method* column which stores the method of optimization that is going to be used (mle, mme, qme, mge), and *log-likelihood* and *aic* columns. 

Before doing anything, we need to call a library called *fitdistrplus* that will do this process for us. After that we can start checking different distributions with the `fitdist` function. 

```{r fitdistrplus}
library(fitdistrplus)
```

# Weibull Distribution
## Weibull with mle
our first method is mle (maximum likelihood estimation):
```{r}
fit_wei_mle = fitdist(time_rep, 'weibull', method='mle')
info_df[1, ] = c('weibull','MLE',fit_wei_mle$loglik, fit_wei_mle$aic)
```

We can use the `summary()` and `plot()` over this fitted weibull distribution to see its characteristics and how it fits the data using plots.

```{r}
summary(fit_wei_mle)
plot(fit_wei_mle)
```

Now we do the same for all other methods and distributions

## Weibull with mme
For the moment matching estimation method we need the moment order of the parameters and a moment function and the `actuar` library for the weibull distribution. 
```{r}
library(actuar)
memp <- function(x, order) mean(x^order)
fit_wei_mme = fitdist(time_rep, 'weibull', method='mme', order=c(1,2), memp=memp)
summary(fit_wei_mme)
plot(fit_wei_mme)
info_df[2, ] = c('weibull','MME',fit_wei_mme$loglik, fit_wei_mme$aic)
```

For the qme method, we need probabilities to check for their quantiles which is stored in the `probs` argument.
## Weibull with qme
```{r}
fit_wei_qme = fitdist(time_rep, 'weibull', method='qme', probs=c(0.5,0.5))
summary(fit_wei_qme)
plot(fit_wei_qme)
info_df[3, ] = c('weibull','QME',fit_wei_qme$loglik, fit_wei_qme$aic)
```

For the maximum goodness-of-fit estimation method we need to add the `gof` argument which stores the goodness-of-fit distance.
## Weibull with gme (KS)
```{r}
fit_wei_mge = fitdist(time_rep, 'weibull', method='mge', gof='KS')
summary(fit_wei_mge)
plot(fit_wei_mge)
info_df[4, ] = c('weibull','MGE KS',fit_wei_mge$loglik, fit_wei_mge$aic)
```


## Weibull with gme (CvM)
```{r}
fit_wei_mge.2 = fitdist(time_rep, 'weibull', method='mge', gof='CvM')
summary(fit_wei_mge.2)
plot(fit_wei_mge.2)
info_df[5, ] = c('weibull','MGE CvM',fit_wei_mge.2$loglik, fit_wei_mge.2$aic)
```


## Weibull with gme (AD)
```{r}
fit_wei_mge.3 = fitdist(time_rep, 'weibull', method='mge', gof='AD')
summary(fit_wei_mge.3)
plot(fit_wei_mge.3)
info_df[6, ] = c('weibull','MGE AD',fit_wei_mge.3$loglik, fit_wei_mge.3$aic)
```

Now we can do this whole process for the other 3 distributions. 

# Gamma Distribution
```{r}
# MLE
fit_gam_mle = fitdist(time_rep, 'gamma', method='mle')
summary(fit_gam_mle)
plot(fit_gam_mle)
info_df[7, ] = c('gamma','MLE',fit_gam_mle$loglik, fit_gam_mle$aic)

# MME
fit_gam_mme = fitdist(time_rep, 'gamma', method='mme')
summary(fit_gam_mme)
plot(fit_gam_mme)
info_df[8, ] = c('gamma','MME',fit_gam_mme$loglik, fit_gam_mme$aic)

# QME
fit_gam_qme = fitdist(time_rep, 'gamma', method='qme', probs=c(0.5,0.5))
summary(fit_gam_qme)
plot(fit_gam_qme)
info_df[9, ] = c('gamma','QME',fit_gam_qme$loglik, fit_gam_qme$aic)

# MGE
fit_gam_mge = fitdist(time_rep, 'gamma', method='mge', gof='KS')
summary(fit_gam_mge)
plot(fit_gam_mge)
info_df[10, ] = c('gamma','MGE KS',fit_gam_mge$loglik, fit_gam_mge$aic)

fit_gam_mge.3 = fitdist(time_rep, 'gamma', method='mge', gof='AD')
summary(fit_gam_mge.3)
plot(fit_gam_mge.3)
info_df[11, ] = c('gamma','MGE AD',fit_gam_mge.3$loglik, fit_gam_mge.3$aic)

```

The only exception here, is that when we use CvM goodness-of-fit distance it cannot perform the optimization process and returns and error. Therefore it is not mentioned here.

# Log-normal Distribution
```{r}
# MLE
fit_lnorm_mle = fitdist(time_rep, 'lnorm', method='mle')
summary(fit_lnorm_mle)
plot(fit_lnorm_mle)
info_df[12, ] = c('lnorm','MLE',fit_lnorm_mle$loglik, fit_lnorm_mle$aic)

# MME
fit_lnorm_mme = fitdist(time_rep, 'lnorm', method='mme')
summary(fit_lnorm_mme)
plot(fit_lnorm_mme)
info_df[13, ] = c('lnorm','MME',fit_lnorm_mme$loglik, fit_lnorm_mme$aic)

# QME
fit_lnorm_qme = fitdist(time_rep, 'lnorm', method='qme', probs=c(0.5,0.5))
summary(fit_lnorm_qme)
plot(fit_lnorm_qme)
info_df[14, ] = c('lnorm','QME',fit_lnorm_qme$loglik, fit_lnorm_qme$aic)

# MGE
fit_lnorm_mge = fitdist(time_rep, 'lnorm', method='mge', gof='KS')
summary(fit_lnorm_mge)
plot(fit_lnorm_mge)
info_df[15, ] = c('lnorm','MGE KS',fit_lnorm_mge$loglik, fit_lnorm_mge$aic)

fit_lnorm_mge.2 = fitdist(time_rep, 'lnorm', method='mge', gof='CvM') 
summary(fit_lnorm_mge.2)
plot(fit_lnorm_mge.2)
info_df[16, ] = c('lnorm','MGE CvM',fit_lnorm_mge.2$loglik, fit_lnorm_mge.2$aic)

fit_lnorm_mge.3 = fitdist(time_rep, 'lnorm', method='mge', gof='AD')
summary(fit_lnorm_mge.3)
plot(fit_lnorm_mge.3)
info_df[17, ] = c('lnorm','MGE AD',fit_lnorm_mge.3$loglik, fit_lnorm_mge.3$aic)
```

# Beta Distribution
For the beta distribution we need to make some adjustments since this distribution needs our data to be within [0-1] interval. We can achieve this by simply dividing our time points to 10000. 

```{r}
time_rep_beta = time_rep/10000
```

Now we can check different methods for this distribution. 
```{r}
# MLE
fit_beta_mle = fitdist(time_rep_beta, 'beta', method='mle')
summary(fit_beta_mle)
plot(fit_beta_mle)
info_df[18, ] = c('beta','MLE',fit_beta_mle$loglik, fit_beta_mle$aic)

# MME
fit_beta_mme = fitdist(time_rep_beta, 'beta', method='mme')
summary(fit_beta_mme)
plot(fit_beta_mme)
info_df[19, ] = c('beta','MME',fit_beta_mme$loglik, fit_beta_mme$aic)

# QME
fit_beta_qme = fitdist(time_rep_beta, 'beta', method='qme', probs=c(0.5,0.5))
summary(fit_beta_qme)
plot(fit_beta_qme)
info_df[20, ] = c('beta','QME',fit_beta_qme$loglik, fit_beta_qme$aic)

# MGE
fit_beta_mge = fitdist(time_rep_beta, 'beta', method='mge', gof='KS')
summary(fit_beta_mge)
plot(fit_beta_mge)
info_df[21, ] = c('beta','MGE KS',fit_beta_mge$loglik, fit_beta_mge$aic)

fit_beta_mge.2 = fitdist(time_rep_beta, 'beta', method='mge', gof='CvM') 
summary(fit_beta_mge.2)
plot(fit_beta_mge.2)
info_df[22, ] = c('beta','MGE CvM',fit_beta_mge.2$loglik, fit_beta_mge.2$aic)

fit_beta_mge.3 = fitdist(time_rep_beta, 'beta', method='mge', gof='AD')
summary(fit_beta_mge.3)
plot(fit_beta_mge.3)
info_df[23, ] = c('beta','MGE AD',fit_beta_mge.3$loglik, fit_beta_mge.3$aic)
```

Now that we have stored all the required information in the `info_df` data frame, we can simply output it to compare different methods and different distributions.

```{r}
info_df
```

To interpret the results, We need to find the distribution with highest log-likelihood value and lowest aic value. As you can see, the best distribution is beta with maximum likelihood estimation method. 

Here you can see this distribution fit again which estimates the parameters 

```{r}
fit_beta_mle
```


The `fitdistrplus` package also has a function called `descdist` that can check different distributions fitness to our data which as you can see gives the same results. 

```{r}
descdist(time_rep)
```


Therefore our chosen model is the beta distribution with parameters `r fit_beta_mle$estimate[1]` as shape1, and `r fit_beta_mle$estimate[2]` as shape2.


## Failure Probability and Reliability
Failure probability and reliability are complimentary of each other, as $R(t) = 1 - F(t)$. Therefore, 5% failure probability equals 95% reliability and vice versa. In order to find the 5% failure probability we can use the `qbeta` function to find the quantile of the optimal beta distribution that produces 5% failure probability or is reliable 95% of the time.

```{r}
fail_quant = qbeta(0.05, fit_beta_mle$estimate[1], fit_beta_mle$estimate[2])
```

Remember that with beta distribution, we divided the time points to 10000. We can now multiply this number to reach the time point that produces 95% reliability. 

Therefore:

```{r}
fail_quant = fail_quant*10000
fail_quant
```

Therefore at `r round(fail_quant,2)` hours, 95% of the systems are still working. 













